### 2021-09-12 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_transfer@e71b02cf58bd95e46d2f4268b1ae53f7f072f07c 🚀

A RDPG estimates the probability of observing interactions between nodes
(species) as a function of the nodes' latent variables. The latent
variables used for the RDPG, called the left and right subspaces, are
defined as \(\mathcal{L} = \mathbf{L}\sqrt{\mathbf{\Sigma}}\), and
\(\mathcal{R} = \sqrt{\mathbf{\Sigma}}\mathbf{R}\) -- using the full
rank of \(\mathbf{A}\), \(\mathcal{L}\mathcal{R}' = \mathbf{A}\), and
using any smaller rank results in
\(\mathcal{L}\mathcal{R}' \approx \mathbf{A}\). Using a rank of 1 for
the t-SVD provides a first-order approximation of the network.

@@

In order to transfer the knowledge from the European metaweb to the
Canadian species pool, we performed ancestral character estimation using
a Brownian motion model, which is a conservative approach in the absence
of strong hypotheses about the nature of phylogenetic signal in the
network decomposition (Litsios \& Salamin 2012). This uses the estimated
feature vectors for the European mammals to create a state
reconstruction for all species (conceptually something akin to a
trait-based mammalian phylogeny using generality and vulnerability
traits) and allows us to impute the missing (latent) trait data for the
Canadian species that are not already in the European network; as we are
focused on predicting contemporary interactions, we only retained the
values for the tips of the tree. We assumed that all traits (\emph{i.e.}
the feature vectors for the left and right subspaces) were independent,
which is a reasonable assumption as every trait/dimension added to the
t-SVD has an \emph{additive} effect to the one before it. Note that the
Upham \emph{et al.} (2019) tree itself has some uncertainty associated
to inner nodes of the phylogeny. In this case study, we have decided to
not propagate this uncertainty, as it would complexify the process. The
Brownian motion algorithm returns the \emph{average} value of the trait,
and its upper and lower bounds. Because we do not estimate other
parameters of the traits' distributions, we considered that every
species trait is represented as a uniform distribution between these
bounds; in a situation where the algorithm would return point values for
all simulations, one could in theory either estimate the parameters of a
distribution for each tip, or draw randomly from the outputs. In all
cases, the inferred left and right sub-spaces for the Canadian species
pool (\(\hat{\mathcal{L}}\) and \(\hat{\mathcal{R}}\)) have entries that
are distributions, representing the range of values for a given species
at a given dimension.

@@

The phylogenetic reconstruction of \(\hat{\mathcal{L}}\) and
\(\hat{\mathcal{R}}\) has an associated uncertainty, represented by the
breadth of the uniform distribution associated to each of their entries.
Therefore, we can use this information to assemble a
\emph{probabilistic} metaweb in the sense of Poisot \emph{et al.}
(2016), \emph{i.e.} in which every interaction is represented as a
single, independent, Bernoulli event of probability \(p\).

@@

Specifically, we have adopted the following approach. For every entry in
\(\hat{\mathcal{L}}\) and \(\hat{\mathcal{R}}\), we draw a value from
its distribution. This results in one instance of the possible left
(\(\hat{\mathcal{l}}\)) and right (\(\hat{\mathcal{r}}\)) subspaces for
the Canadian metaweb. These can be multiplied, to produce one matrix of
real values. Because the entries in \(\hat{\mathcal{l}}\) and
\(\hat{\mathcal{r}}\) are in the same space where \(\mathcal{L}\) and
\(\mathcal{R}\) were originally predicted, it follows that the threshold
\(\rho\) estimated for the European metaweb also applies. We use this
information to produce one random Canadian metaweb,
\(N = \hat{\mathcal{L}}\)\(\hat{\mathcal{R}}' \ge \rho\). As we can see
in (fig.~\ref{fig:subspaces}) the European and Canadian metawebs are
structurally similar (as would be expected given the biogeographic
similarities) and the two (left and right) subspaces are distinct
\emph{i.e.} capturing predation (generality) and prey (vulnerability)
traits.

>>>>

A RDPG estimates the probability of observing interactions between nodes
(species) as a function of the nodes' latent variables. The latent
variables used for the RDPG, called the left and right subspaces, are
defined as \(\mathscr{L} = \mathbf{L}\sqrt{\mathbf{\Sigma}}\), and
\(\mathscr{R} = \sqrt{\mathbf{\Sigma}}\mathbf{R}\) -- using the full
rank of \(\mathbf{A}\), \(\mathscr{L}\mathscr{R}' = \mathbf{A}\), and
using any smaller rank results in
\(\mathscr{L}\mathscr{R}' \approx \mathbf{A}\). Using a rank of 1 for
the t-SVD provides a first-order approximation of the network.

@@

In order to transfer the knowledge from the European metaweb to the
Canadian species pool, we performed ancestral character estimation using
a Brownian motion model, which is a conservative approach in the absence
of strong hypotheses about the nature of phylogenetic signal in the
network decomposition (Litsios \& Salamin 2012). This uses the estimated
feature vectors for the European mammals to create a state
reconstruction for all species (conceptually something akin to a
trait-based mammalian phylogeny using generality and vulnerability
traits) and allows us to impute the missing (latent) trait data for the
Canadian species that are not already in the European network; as we are
focused on predicting contemporary interactions, we only retained the
values for the tips of the tree. We assumed that all traits (\emph{i.e.}
the feature vectors for the left and right subspaces) were independent,
which is a reasonable assumption as every trait/dimension added to the
t-SVD has an \emph{additive} effect to the one before it. Note that the
Upham \emph{et al.} (2019) tree itself has some uncertainty associated
to inner nodes of the phylogeny. In this case study, we have decided to
not propagate this uncertainty, as it would complexify the process. The
Brownian motion algorithm returns the \emph{average} value of the trait,
and its upper and lower bounds. Because we do not estimate other
parameters of the traits' distributions, we considered that every
species trait is represented as a uniform distribution between these
bounds; in a situation where the algorithm would return point values for
all simulations, one could in theory either estimate the parameters of a
distribution for each tip, or draw randomly from the outputs. In all
cases, the inferred left and right sub-spaces for the Canadian species
pool (\(\hat{\mathscr{L}}\) and \(\hat{\mathscr{R}}\)) have entries that
are distributions, representing the range of values for a given species
at a given dimension.

@@

The phylogenetic reconstruction of \(\hat{\mathscr{L}}\) and
\(\hat{\mathscr{R}}\) has an associated uncertainty, represented by the
breadth of the uniform distribution associated to each of their entries.
Therefore, we can use this information to assemble a
\emph{probabilistic} metaweb in the sense of Poisot \emph{et al.}
(2016), \emph{i.e.} in which every interaction is represented as a
single, independent, Bernoulli event of probability \(p\).

@@

Specifically, we have adopted the following approach. For every entry in
\(\hat{\mathscr{L}}\) and \(\hat{\mathscr{R}}\), we draw a value from
its distribution. This results in one instance of the possible left
(\(\hat{\mathscr{l}}\)) and right (\(\hat{\mathscr{r}}\)) subspaces for
the Canadian metaweb. These can be multiplied, to produce one matrix of
real values. Because the entries in \(\hat{\mathscr{l}}\) and
\(\hat{\mathscr{r}}\) are in the same space where \(\mathscr{L}\) and
\(\mathscr{R}\) were originally predicted, it follows that the threshold
\(\rho\) estimated for the European metaweb also applies. We use this
information to produce one random Canadian metaweb,
\(N = \hat{\mathscr{L}}\)\(\hat{\mathscr{R}}' \ge \rho\). As we can see
in (fig.~\ref{fig:subspaces}) the European and Canadian metawebs are
structurally similar (as would be expected given the biogeographic
similarities) and the two (left and right) subspaces are distinct
\emph{i.e.} capturing predation (generality) and prey (vulnerability)
traits.
